---
title: "lab14_hw"
output: 
  html_document: 
    keep_md: yes
date: '2022-02-28'
---



### Libraries

### Load the Libraries
```{r}
library(tidyverse)
library(janitor)
library("palmerpenguins")
#install.packages("ggVennDiagram")
library(ggVennDiagram)
library(RColorBrewer)
#install.packages("ggworldcloud")
library(ggwordcloud)
options(scipen=999) #cancels the use of scientific notation for the session
```


### Data
```{r}
superhero_info <- readr::read_csv("data/heroes_information.csv", na = c("", "-99", "-"))
superhero_powers <- readr::read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))
beachbugs_long <- readr::read_csv("data/beachbugs_long.csv")
```



1. Clean up the column names (no capitals, not spaces) of `superhero_info`, then use 2 functions to remind yourself of structure of the `superhero_info` data set.

```{r}
superhero_info <- readr::read_csv("data/heroes_information.csv", na = c("", "-99", "-"))%>%clean_names()
superhero_powers <- readr::read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))%>%clean_names()
```

```{r}
beachbugs_long <- readr::read_csv("data/beachbugs_long.csv")%>%clean_names()
```
```{r}
glimpse(superhero_powers)
```
```{r}
summary(superhero_powers)
```
```{r}
glimpse(superhero_info)
```
```{r}
summary(superhero_info)
```
```{r}
glimpse(beachbugs_long)
```
```{r}
summary(beachbugs_long)
```

```{r}
superhero_powers
```
```{r}
superhero_info
```
```{r}
beachbugs_long
```

2. Are bad guys bigger? Make box-plots of weight by `alignment` (alignment on the x-axis).

```{r}
superhero_info%>%
  filter(!weight=="NA", !alignment=="NA")%>%
  ggplot(aes(x=alignment, y=weight,fill=alignment))+
  geom_boxplot()
```

3. Now, make a violin plot of weight by `alignment` (alignment on the x-axis). Add some color! What information can you observe in the violin plot that was not visible in the boxplot?

```{r}
superhero_info%>%
  filter(!weight=="NA", !alignment=="NA")%>%
  ggplot(aes(x=alignment, y=weight,fill=alignment))+
  geom_violin()
```


4. Use `alpha = .5` in `geom_boxplot()` and `geom_violin()` to layer both plots on top of one another. What does this tell you about the distribution of weight in "`bad`" guys?

```{r}
superhero_info%>%
  filter(!weight=="NA", !alignment=="NA")%>%
  ggplot(aes(x=alignment, y=weight,fill=alignment))+
  geom_boxplot(color = "white", alpha = .3, na.rm = TRUE) +
  geom_violin(alpha = .6, na.rm = TRUE)
```
#This shows us the actual distribuition of weight and where alignments are more found on average compared to box plots.

5. Box plots are great for showing how the distribution of a numeric variable (e.g. weight) varies among a categorical variable (e.g. alignment).
  Make your own violin plot with the superhero data. 
  What is your numeric data? Height
  What is your categorical variable? Alignment

```{r}
superhero_info%>%
  filter(!height=="NA", !alignment=="NA")%>%
  ggplot(aes(x=alignment, y=height,fill=alignment))+
  geom_boxplot(color = "white", alpha = .3, na.rm = TRUE) +
  geom_violin(alpha = .6, na.rm = TRUE)
```

6. Remind yourself what `beachbugs` looks like. Then generate a heatmap of buglevels by site and year. 
color it with `scale_fill_gradient(low="yellow", high="red")` or colors of your choice. you may find it looks best with one color!
(dont forget, `coord_flip()` is a quick way to improve the look of your plot if you dont like the default orientation)

```{r}
beachbugs_long
```
```{r}
beachbugs_long%>%
  ggplot(aes(year, site, fill=buglevels))+
  geom_tile()+
  scale_fill_gradient(low = "navyblue", high = "magenta")
```

7. Use the provided code to normalize the beachbug data set. 
Then make a heatmap with the `beachbugs_normalized` data, and use the same color choice as above. Which heatmap do you think is more informative? why?

```{r}
beachbugs_normalized<-beachbugs_long%>%
  group_by(site)%>%
  arrange(year, site)%>%
  mutate(max_buglevel = max(buglevels, na.rm=T)) %>% 
  mutate(buglevel=buglevels/max_buglevel)%>%
  select(year, site, buglevel)
beachbugs_normalized
```
```{r}
beachbugs_normalized%>%
  ggplot(aes(year, site, fill=buglevel))+
  geom_tile()+
  scale_fill_gradient(low = "navyblue", high = "magenta")
```

#The normalized version is more helpful to interpret because it showcases buglevel for each beach in during a certain time period. I don't think that makes sense. 

8. Let's make a venn diagram of `superhero_info`, from 4 questions:
Is their alignment evil? 

```{r}
superhero_info
```
```{r}
yellow_eye<-superhero_info%>%
  filter(eye_color=="yellow")%>%
  pull(name)
yellow_eye
```
```{r}
animal_race<-superhero_info%>%
  filter(race=="Mutant")%>%
  pull(name)
animal_race
```

```{r}
hair<-superhero_info%>%
  filter(hair_color=="White")%>%
  pull(name)
hair
```

```{r}
evil <- superhero_info %>%
filter(alignment == "bad")%>%
pull(name)
evil
```

9. Let's make the venn diagram! use the code from lab as a reference. 

```{r}
evil_diagram <- list(evil, hair, animal_race, yellow_eye)

ggVennDiagram(evil_diagram, category.names = c("Evil", "White Hair", "Mutant", "Yellow Eyes"))
```

10. Choose one intersection of the venn diagram that is interesting to you. Use dplyr to find the names of the superheros in that intersection.

```{r}
superhero_info%>%
  filter(alignment=="bad")%>%
  filter(hair_color=="White")%>%
select(name)
```

11. Make another venn diagram with the `superhero_info` data. What are your questions? ( At least 2!) 

-Is their alignment neutral?
-Are they male?
-Do they have red eyes?

```{r}
neutral <- superhero_info %>%
filter(alignment == "neutral")%>%
pull(name)
neutral
```

```{r}
gender<-superhero_info%>%
  filter(gender=="Male")%>%
  pull(name)
gender
```

```{r}
red_eyes<-superhero_info%>%
  filter(eye_color=="red")%>%
  pull(name)
red_eyes
```

```{r}
neutral_diagram <- list(neutral, gender, red_eyes)

ggVennDiagram(neutral_diagram, category.names = c("Neutral", "Male", "Red Eyes"))+
  scale_fill_distiller(palette = "green")
```
```{r}
eleven_list<-superhero_info%>%
  filter(alignment == "neutral"& gender=="Male"& eye_color=="red")
eleven_list
```

12.What are some very common super powers? Lets make a word cloud with the `superhero_powers` data.

```{r}
super_hero_freq<-superhero_powers%>%
  select(-hero_names)%>%
  summarise_all(~sum(.))%>%
  pivot_longer(everything(),
               names_to = "power",
               values_to = "frequency")%>%
  top_n(50)%>%
  arrange(desc(frequency))
super_hero_freq
```
```{r}
super_hero_freq%>%
  ggplot(aes(
    size=frequency,
    color=power,
    label=power))+
  geom_text_wordcloud()+
  scale_size_area(max_size = 8)
```

13.Who are some very powerful supers? Lets make a different word cloud with the `superhero_powers` data. Use the provided code to make the frequency table, then make a word cloud with it. You can use `hero_names` for the labels, and `sum_powers` for size.

```{r}
power_quantity <- superhero_powers %>% 
  pivot_longer(-hero_names, names_to = "power", values_to = "yes_or_no") %>% 
  group_by(hero_names) %>% 
  mutate(sum_powers = sum(yes_or_no, na.rm=T)) %>%
  arrange(desc(sum_powers), hero_names, desc(yes_or_no))
power_quantity
#we have to trim down to only the top 50, or it will crowd the image!
power_quantity <- power_quantity %>% 
  ungroup %>%
  distinct(hero_names, sum_powers) %>%
  slice_max(sum_powers, n = 50)
power_quantity
```

```{r}
power_quantity%>%
  ggplot(aes(
    size=sum_powers,
    color=hero_names,
    label=hero_names))+
  geom_text_wordcloud()+
  scale_size_area(max_size = 5)
```


